---
title: "Kepler Candidates"
author: "J. Dotson"
date: "8/23/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

require(ggplot2)
require(hexbin)
require(scales)
require(dplyr)
require(tidyverse)
require(readr)
require(RColorBrewer)
require(ggforce)
require(ggthemes)
require(gridExtra)
myecho=FALSE
knitr::opts_knit$set(root.dir = '../.')
```


```{r}
download <- FALSE
source("Rscripts/mycolors.R")
```


```{r }

baseurl<-"https://exoplanetarchive.ipac.caltech.edu/cgi-bin/nstedAPI/nph-nstedAPI?table=cumulative"


params <- "kepid,kepoi_name,kepler_name,koi_disposition,koi_vet_stat,koi_vet_date,koi_pdisposition,koi_score,koi_disp_prov,koi_period,koi_eccen,koi_impact,koi_duration,koi_prad,koi_sma,koi_incl,koi_teq,koi_insol,koi_model_snr,koi_steff,koi_slogg,koi_smet,koi_srad,koi_smass,koi_sage,koi_kepmag"



address<-paste(baseurl,"&select=",params,sep="")

if (download){
  download.file(address, "data/cumulative_kois.csv")
  print(paste("Cumulative KOI table downloaded on ",date()))
}


```


```{r }
kois <- read_csv("data/cumulative_kois.csv")
```

Let's just keep the one that Kepler has ordained as a candidate...

```{r}

count(kois,koi_pdisposition) #project's disposition
count(kois,koi_disposition)  #archive's disposition

candidates <- kois %>% filter(koi_pdisposition == "CANDIDATE")

```

```{r}
count(candidates,koi_disposition)

#let's just get rid of that one false positive in the nexsci disposition for plotting clarity...
candidates <- candidates %>% filter(koi_disposition != "FALSE POSITIVE")
```

Note -- the confirmed planet count here is less than the active confirmed planet count at NExScI.   But we're not going to stress about that.  (The number of  candidates is the same...  as of August 23, 2019)


One last thing before we start plotting...  Let's assign them to a size category.  Using the same categories usec elsewhere...

Earth-size:  < 1.25 Re
Super earth size: 1.25 - 2Re
Neptune size: 2 - 6 Re
Jupiter size: 6 - 15 Re
larger: 15 - 25 Re


```{r}

earth_max <- 1.25
super_earth_max <- 2.0
neptune_max <- 6.0
jupiter_max <- 15.0

candidates$size <- rep(NA, nrow(candidates))

candidates <- mutate(candidates, size = ifelse(((koi_prad > neptune_max) & (koi_prad <= jupiter_max)), "Jupiter",
                                             ifelse(((koi_prad > super_earth_max) & (koi_prad <= neptune_max)), "Neptune",
                                                    ifelse(((koi_prad > earth_max)& (koi_prad <= super_earth_max)), "SuperEarth",
                                                           ifelse(((koi_prad > 0) & (koi_prad <= earth_max)), "Earth", "Huge")))))

count(candidates, size)



```
It looks like all the candidates without radii are from a supplementary delivery.  I'm ok with dumping these for plotting...

```{r}
candidates <- candidates %>% filter(!is.na(size))
candidates$size <- factor(candidates$size, levels=c("Earth","SuperEarth","Neptune","Jupiter","Huge"))
```




```{r}

baseplot <- ggplot(candidates) + theme_bw()

myylim<-c(0.3,25)
myxlim<-c(0.1,1000)

plot1 <- baseplot + geom_point(aes(x=koi_period,y=koi_prad,col=koi_disposition),alpha=0.5)
plot1 <- plot1 +  scale_x_log10(limits=myxlim,breaks=c(1,10,100,1000)) + 
  scale_y_log10(limits=myylim,breaks=c(1,4,10,20,40))
plot1 <- plot1 +  ylab("Size Relative to Earth (Radius)") +
  xlab("Orbital Period (Days)") + theme(legend.title = element_blank())

print(plot1)

plot1 <- baseplot + geom_point(aes(x=koi_period,y=koi_prad),alpha=0.5,col="cadetblue")
plot1 <- plot1 +  scale_x_log10(limits=myxlim,breaks=c(1,10,100,1000)) + 
  scale_y_log10(limits=myylim,breaks=c(1,4,10,20,40))
plot1 <- plot1 +  ylab("Size Relative to Earth (Radius)") +
  xlab("Orbital Period (Days)") 
print(plot1)

plot2 <- baseplot + geom_hex(aes(x=koi_period,y=koi_prad),bins=70) + scale_fill_continuous(type="viridis")
plot2 <- plot2 +  scale_x_log10(limits=myxlim,breaks=c(1,10,100,1000)) + 
  scale_y_log10(limits=myylim,breaks=c(1,4,10,20,40))
plot2 <- plot2 +  ylab("Size Relative to Earth (Radius)") +
  xlab("Orbital Period (Days)") 
print(plot2)

plot2 <- baseplot + geom_bin2d(aes(x=koi_period,y=koi_prad),bins=70) + scale_fill_continuous(type="viridis")
plot2 <- plot2 +  scale_x_log10(limits=myxlim,breaks=c(1,10,100,1000)) + 
  scale_y_log10(limits=myylim,breaks=c(1,4,10,20,40))
plot2 <- plot2 +  ylab("Size Relative to Earth (Radius)") +
  xlab("Orbital Period (Days)") 
print(plot2)


```

```{r}
# this function is necessary to plot axes that are both reversed and log

reverselog_trans <- function(base = exp(1)) {
    trans <- function(x) -log(x, base)
    inv <- function(x) base^(-x)
    trans_new(paste0("reverselog-", format(base)), trans, inv, 
              log_breaks(base = base), 
              domain = c(1e-100, Inf))
}

```




```{r}
plot3 <- baseplot + geom_point(aes(x=koi_insol,y=koi_steff,col=size),alpha = 0.5) + scale_color_geert(palette="planets")
plot3 <- plot3 +  scale_y_log10() + scale_x_continuous(trans=reverselog_trans(10))
plot3 <- plot3 + xlab("stellar flux (earth units)") + ylab("Stellar Effective Temperature (K)")
#plot3 <- plot3 + scale_color_hue(breaks=c("Earth","SuperEarth","Neptune","Jupiter","Huge"))
print(plot3)

plot3 <- plot3 + facet_wrap(~size)
print(plot3)
```

```{r}
calc_Seff <- function(Teff){
  criteria <- c("recent_venus", "runaway_greenhouse_1", "runaway_greenhouse_5", "runaway_greenhouse_01","max_greenouse","early_mars" )
  Ssol <- c(1.776,1.107,1.188,0.99,0.356,0.32)
  a <- c(2.136e-4,1.332e-4,1.433e-4,1.209e-4,6.171e-5,5.547e-5)
  b <- c(2.533e-8,1.58e-8,1.707e-8,1.404e-8,1.698e-9,1.526e-9)
  c <- c(-1.332e-11,-8.308e-12,-8.968e-12,-7.418e-12,-3.198e-12,-2.874e-12)
  d <- c(-3.097e-15,-1.931e-15,-2.084e-15,-1.713e-15,-5.575e-16,-5.011e-16)
  HZ_def <- data.frame(criteria,Ssol,a,b,c,d)
  
  Tst <- Teff - 5780 #assumes Teff is in Kelvin
  Seff <- HZ_def$Ssol + HZ_def$a*Tst + HZ_def$b*Tst^2 + HZ_def$c*Tst^3 + HZ_def$d*Tst^4
  result <- data.frame(criteria, Seff)
  return(result)
}
```


I suspect there's a more elegant way to do this, but...  Kinda don't want to reinvent the wheel!

```{r}

temperaturelist <- seq(from=2600, to=7200, by=100)

my_HZ_limits <- tibble(T=temperaturelist)
my_HZ_limits$recent_venus <- rep(NA, nrow(my_HZ_limits))
my_HZ_limits$early_mars <- rep(NA, nrow(my_HZ_limits))

for (i in 1:length(temperaturelist)){
  temp <- calc_Seff(my_HZ_limits$T[i])
  jj <- which(temp$criteria == "recent_venus")
  my_HZ_limits$recent_venus[i] <- temp$Seff[jj]
  jj <- which(temp$criteria == "early_mars")
  my_HZ_limits$early_mars[i] <- temp$Seff[jj]
  
}

templot <- ggplot(my_HZ_limits) + theme_bw()
templot <- templot + geom_point(aes(x=recent_venus,y=T),col="red")
templot <- templot + geom_point(aes(x=early_mars,y=T),col="blue")
templot <- templot + scale_y_log10() + scale_x_continuous(trans=reverselog_trans(10))
print(templot)

#probably need to reorganize this data...

long_HZ_limits <- gather(my_HZ_limits, criteria, Seff, recent_venus:early_mars, factor_key=TRUE)
#newdata <- mtcars[order(mpg, cyl),]
long_HZ_limits <- long_HZ_limits[order(long_HZ_limits$T, long_HZ_limits$Seff),]

templot <- ggplot(long_HZ_limits) + theme_bw()
templot <- templot + geom_polygon(aes(x=Seff,y=T),col="red")
templot <- templot + scale_y_log10() + scale_x_continuous(trans=reverselog_trans(10))
print(templot)


#+ coord_flip()


templot <- ggplot(my_HZ_limits) + theme_bw()
templot <- templot + geom_ribbon(aes(x=T, ymin=recent_venus, ymax=early_mars),fill="lightgrey")
#templot <- templot + scale_y_continuous(trans=reverselog_trans(10)) + scale_y_log10()
templot <- templot + coord_flip()

print(templot)
templot <- templot + scale_x_log10() + scale_y_continuous(trans=reverselog_trans(10))
print(templot)

HZrange <- templot

```


```{r}


templot <- ggplot(my_HZ_limits) + theme_bw() + scale_color_geert(palette="planets2")
templot <- templot + geom_ribbon(aes(x=T, ymin=recent_venus, ymax=early_mars),fill="lightgrey")
templot <- templot + geom_point(data=candidates, aes(x=koi_steff, y=koi_insol, col=size), alpha=0.8, size=0.7)
print(templot)

templot <- templot + coord_flip()
print(templot)

templot <- templot + scale_x_log10() + scale_y_continuous(trans=reverselog_trans(10)) 
templot <- templot + theme(axis.title = element_text(size=14)) + theme(legend.title = element_text(size=14))
templot <- templot + theme(axis.text = element_text(size=12)) + theme(legend.text = element_text(size=12))
templot <- templot + xlab("Stellar Effective Temperature (K)") + ylab("Stellar Flux (earth units)")
templot <- templot + guides(colour = guide_legend(override.aes = list(size=3)))
print(templot)

templot2 <- templot + facet_wrap(~size)
print(templot2)

```






