---
title: "Animated Planets"
author: "J. Dotson"
date: "6/15/2019"
output: html_document
---

Animated plots showing evolution of confirmed planets over time...


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

require(ggplot2)
require(dplyr)
require(tidyverse)
require(readr)
require(RColorBrewer)
require(ggforce)
require(ggthemes)
require(gridExtra)
require(gganimate)
require(tidyr)

myecho=FALSE
knitr::opts_knit$set(root.dir = '../.')
```

Using the planet property table created in planethistory_base.  

```{r }

p_planets <- readRDS("RDS/planets_to_plot.rds")
source("Rscripts/mycolors.R")

```

T
```{r}
p_planets$facility <- rep("All Other Discoveries", nrow(p_planets))
jj <- which(p_planets$pl_instrument == "Kepler CCD Array")
p_planets$facility[jj] <- "Kepler/K2 Discoveries"

p_planets$facility <- factor(p_planets$facility, levels=c("Kepler/K2 Discoveries","All Other Discoveries"))


count(p_planets, facility)
```


here are a couple of years for which no planets are discovered.  Lets add some empty planets for those years.

```{r}
count(p_planets, pl_disc)

# need 1990, 1991, and 1993
orig_p_planets <- p_planets

p_planets <- add_row(p_planets,pl_disc=1990, facility="All Other Discoveries")
p_planets <- add_row(p_planets,pl_disc=1991, facility="All Other Discoveries")
p_planets <- add_row(p_planets,pl_disc=1993, facility="All Other Discoveries")


```



Starting with our usual scatter plot...

```{r }

baseplot<-ggplot(p_planets) + theme_bw() +  ylab("Size Relative to Earth (Radius)") +
  xlab("Orbital Period (Days)")

myylim<-c(0.3,25)
myxlim<-c(0.1,1000)

baseplot2 <- baseplot + scale_x_log10(limits=myxlim,breaks=c(1,10,100,1000)) + 
  scale_y_log10(limits=myylim,breaks=c(1,4,10,20,40))

thisplot <- baseplot2 + geom_point(aes(x=pl_orbper,y=new_rade,group=seq_along(pl_disc)),
                                   shape=21,fill=geert_cols("wisteria"),col="white",size=4)
#
print(thisplot)

thisplot <- thisplot + transition_time(pl_disc) + 
  labs(title="Confirmed Planets as of {floor(frame_time)}")
thisplot <- thisplot + shadow_mark(past=TRUE, col="white",fill=geert_cols("silver")) + enter_appear()
thisplot <- thisplot + theme(plot.title=element_text(size=16),
                             axis.title.x=element_text(size=14),
                             axis.title.y=element_text(size=14),
                             axis.text = element_text(size=12))


# this call to animate will make an mp4
#animate(thisplot, nframes = 300,renderer = av_renderer(),width=1200, height=800, res=144)
animate(thisplot, nframes = 30, fps=1, end_pause=5,renderer = av_renderer(),width=1200, height=750, res=144)

# this call to animate will make a gif
#animate(thisplot, nframes=300, width=1200, height=800, res=144) 
#animate(thisplot,nframes=300)




```


```{r}
anim_save("plots/scatter_history.gif")
```

Let's make a version that highlights the Kepler planets.  First let's make a Kepler/not Kepler column to use as a key.


```{r}
baseplot<-ggplot(p_planets) + theme_bw() +  ylab("Size Relative to Earth (Radius)") +
  xlab("Orbital Period (Days)")

myylim<-c(0.3,25)
myxlim<-c(0.1,1000)

baseplot2 <- baseplot + scale_x_log10(limits=myxlim,breaks=c(1,10,100,1000)) + 
  scale_y_log10(limits=myylim,breaks=c(1,4,10,20,40))

thisplot <- baseplot2 + geom_point(aes(x=pl_orbper,y=new_rade,group=seq_along(pl_disc),
                                       fill=facility),
                                   shape=21,col="white",size=4)
#
print(thisplot)

thisplot <- thisplot + transition_time(pl_disc) + labs(title="Year: {floor(frame_time)}")
thisplot <- thisplot + shadow_mark(past=TRUE, col="white",fill="darkgrey") + enter_appear()
thisplot <- thisplot + theme(plot.title=element_text(size=16),
                             axis.title.x=element_text(size=14),
                             axis.title.y=element_text(size=14))


# this call to animate will make an mp4
#animate(thisplot, nframes = 300,renderer = av_renderer(),width=1200, height=800, res=144)

# this call to animate will make a gif
#animate(thisplot, nframes=300, width=1200, height=800, res=144) 
animate(thisplot,nframes=300)

```



It looks like gganimate has problems when a facet doesn't have data in a given transition step.  So, need to add some empty points for a few years.  But which...

```{r}

temp <- p_planets[jj,]

count(temp,pl_disc)


```

So, let's add empty planets for Kepler for every year between 1989 and 2008 (inclusive)

```{r}
p_planets <- add_row(p_planets, pl_disc=1989:2008, facility="Kepler/K2 Discoveries")
```


Would like to label the number of planets...
```{r}


mycum <- function(df){
  df$cum <- rep(0, nrow(df))
  df$cum[1] <- df$n[1]
  for (i in 2:nrow(df)){
    df$cum[i] <- df$n[i] + df$cum[i-1]
  }
  return(df)
}


p_planets2 <- p_planets
p_planets2$count <- rep(NA,nrow(p_planets2))

jj <- which(orig_p_planets$facility == "Kepler/K2 Discoveries")
kepcount <- count(orig_p_planets[jj,], pl_disc)
kepcount <- mycum(kepcount)
p_planets2 <- add_row(p_planets2, pl_disc=1989:2008,facility="Kepler/K2 Discoveries", count=0)
p_planets2 <- add_row(p_planets2, pl_disc=kepcount$pl_disc,facility="Kepler/K2 Discoveries",count=kepcount$cum)

jj <- which(orig_p_planets$facility == "All Other Discoveries")
notcount <- count(orig_p_planets[jj,], pl_disc)
notcount <- mycum(notcount)
p_planets2 <- add_row(p_planets2, pl_disc=1990:1991, facility="All Other Discoveries",count=1)
p_planets2 <- add_row(p_planets2, pl_disc=1993, facility="All Other Discoveries", count=3)
p_planets2 <- add_row(p_planets2, pl_disc=notcount$pl_disc, facility="All Other Discoveries", count=notcount$cum)


```



```{r}
baseplot<-ggplot(p_planets2) + theme_bw() +  ylab("Size Relative to Earth (Radius)") +
  xlab("Orbital Period (Days)")

myylim<-c(0.3,25)
myxlim<-c(0.1,1000)

baseplot2 <- baseplot + scale_x_log10(limits=myxlim,breaks=c(1,10,100,1000)) + 
  scale_y_log10(limits=myylim,breaks=c(1,4,10,20,40))

thisplot <- baseplot2 + geom_point(aes(x=pl_orbper,y=new_rade,group=seq_along(pl_disc)),
                                   shape=21,fill=geert_cols("wisteria"),col="white",size=4)
thisplot <- thisplot + facet_grid(aes(cols=facility))
#
print(thisplot)

thisplot <- thisplot + transition_time(pl_disc) + 
  labs(title="Confirmed Planets as of {floor(frame_time)}")
thisplot <- thisplot + shadow_mark(past=TRUE, col="white",fill=geert_cols("silver"),exclude_layer=2) 
thisplot <- thisplot + enter_appear()
thisplot <- thisplot + theme(plot.title=element_text(size=18, hjust=0.5),
                             axis.title.x=element_text(size=14),
                             axis.title.y=element_text(size=14),
                             strip.text.x = element_text(size = 14),
                             axis.text = element_text(size=12))
thisplot <- thisplot + geom_label(aes(label=floor(count),x=100,y=0.5), 
                                  size=10, fill="white",
                                  linetype=0)
#thisplot <- thisplot + shadow_mark(past=TRUE, exclude_layer=1)


# this call to animate will make an mp4
animate(thisplot, nframes = 30, fps=1, end_pause=5,renderer = av_renderer(),width=1200, height=750, res=144)

# this call to animate will make a gif
#animate(thisplot, nframes=300, width=1200, height=800, res=144) 
#animate(thisplot,nframes=30,fps=1, end_pause=5, width=800, height=500)
```

```{r}
anim_save("plots/compare_scatter_history.gif")
```


Let's sort the planets into size bins...

Using the bins we've used in our public materials (btw, NExScI uses these bins also...)

Earth-size:  < 1.25 Re
Super earth size: 1.25 - 2Re
Neptune size: 2 - 6 Re
Jupiter size: 6 - 15 Re
larger: 15 - 25 Re


mutate(gradebook, letter = ifelse(grade %in% 60:69, "D",
                                     ifelse(grade %in% 70:79, "C",
                                            ifelse(grade %in% 80:89, "B",
                                                   ifelse(grade %in% 90:99, "A", "F")))))

```{r}

earth_max <- 1.25
super_earth_max <- 2.0
neptune_max <- 6.0
jupiter_max <- 15.0

p_planets <- orig_p_planets

p_planets$size <- rep(NA, nrow(p_planets))

p_planets <- mutate(p_planets, size = ifelse(((new_rade > neptune_max) & (new_rade <= jupiter_max)), "Jupiter",
                                             ifelse(((new_rade > super_earth_max) & (new_rade <= neptune_max)), "Neptune",
                                                    ifelse(((new_rade > earth_max)& (new_rade <= super_earth_max)), "SuperEarth",
                                                           ifelse(((new_rade > 0) & (new_rade <= earth_max)), "Earth", "Huge")))))


```




```{r}

#just for now, remove the planets without a radius...
rad_planets <- p_planets %>% filter(!is.na(size))

rad_planets$size <- factor(rad_planets$size, levels=c("Earth","SuperEarth","Neptune","Jupiter","Huge"))

baseplot <- ggplot(rad_planets) + theme_bw()

bar_p <- baseplot + geom_bar(aes(size,col=size, fill=size)) 
bar_p <- bar_p + scale_fill_geert(palette="planets2") + scale_color_geert(palette="planets2")

bar_p<- bar_p + xlab("") + ylab("Number Known")
bar_p <- bar_p + theme(legend.position = "none")
bar_p <- bar_p + theme(axis.text=element_text(size=14))
bar_p <- bar_p + theme(axis.title=element_text(size=16))
bar_p <- bar_p + theme(plot.title=element_text(size=20, face="bold",hjust=0.5))


print(bar_p)
```



To animate, we need to reorganize the data bit so that the cumulative counts per year are available. 

Need column for count, size, and year.  

First figure out how many of each size were discovered in a given year.


```{r}
temp <- rad_planets %>% filter(size == "Earth")
earth_history <- count(temp, pl_disc)
colnames(earth_history) <- c("year", "Earths")

temp <- rad_planets %>% filter(size == "SuperEarth")
sd_history <- count(temp, pl_disc)
colnames(sd_history) <- c("year", "SuperEarths")

temp <- rad_planets %>% filter(size == "Neptune")
nep_history <- count(temp, pl_disc)
colnames(nep_history) <- c("year", "Neptunes")

temp <- rad_planets %>% filter(size == "Jupiter")
jup_history <- count(temp, pl_disc)
colnames(jup_history) <- c("year","Jupiters")

temp <- rad_planets %>% filter(size == "Huge")
huge_history <- count(temp, pl_disc)
colnames(huge_history) <- c("year", "HugePlanets")

disc_history <- count(rad_planets, pl_disc)
colnames(disc_history) <- c("year", "Planets")

size_history <- merge(disc_history, earth_history, by.x="year", by.y="year", all.x=TRUE)
size_history <- merge(size_history, sd_history, by.x="year",by.y="year",all.x=TRUE)
size_history <- merge(size_history, nep_history, by.x="year", by.y="year",all.x=TRUE)
size_history <- merge(size_history, jup_history, by.x="year", by.y="year", all.x=TRUE)
size_history <- merge(size_history, huge_history, by.x="year", by.y = "year", all.x=TRUE)




```

Next calculate the cumulative number in each size bin as a function of time.

```{r}
cum_column <- function(myx, myy){
  myresult <- tibble(x = myx, y=myy)
  jj <- which(is.na(myresult$y))
  myresult$y[jj] <- 0
  myresult$cum <- rep(0,nrow(myresult))
  myresult$cum[1] <- myresult$y[1]
  for (i in 2:length(myx)){
    myresult$cum[i] <- myresult$cum[i-1] + myresult$y[i]
  }
  return(myresult$cum)
}
```


```{r}
size_history$earth_cum <- cum_column(size_history$year, size_history$Earths)
size_history$sd_cum <- cum_column(size_history$year, size_history$SuperEarths)
size_history$nep_cum <- cum_column(size_history$year, size_history$Neptunes)
size_history$jup_cum <- cum_column(size_history$year, size_history$Jupiters)
size_history$huge_cum <- cum_column(size_history$year, size_history$HugePlanets)
```


And now put the cumulative counts into a tidy long format
```{r}

temp <- size_history %>% select(c("year", "earth_cum", "sd_cum","nep_cum","jup_cum","huge_cum"))
colnames(temp) <- c("year","earth","superearth","neptune","jupiter","huge")

size_history_long <- gather(temp, size, count, earth:huge)


size_history_long$size_f <- factor(size_history_long$size, levels=c("earth","superearth","neptune","jupiter","huge"))

```


Finally plot!
```{r}

temp <- size_history_long %>% filter(year == 2019)

baseplot <- ggplot(temp) + theme_bw()

p1 <- baseplot + geom_bar(aes(x=size_f, y=count, col=size_f,fill=size_f),stat="identity")
p1 <- p1 + scale_color_geert(palette="planets2") + scale_fill_geert(palette="planets2")
p1<- p1 + xlab("size") + ylab("Number Known")
p1 <- p1 + theme(legend.position = "none")
p1 <- p1 + theme(axis.text=element_text(size=14))
p1 <- p1 + theme(axis.title=element_text(size=16))
p1 <- p1 + theme(plot.title=element_text(size=20, face="bold",hjust=0.5))
print(p1)
```

And animate...


```{r}
baseplot <- ggplot(size_history_long) + theme_bw()

p1 <- baseplot + geom_bar(aes(x=size_f, y=count, col=size_f,fill=size_f),stat="identity")
p1 <- p1 + scale_color_geert(palette="planets2") + scale_fill_geert(palette="planets2")
p1 <- p1 + xlab("size") + ylab("Number Known")
p1 <- p1 + theme(legend.position = "none")
p1 <- p1 + theme(axis.text=element_text(size=14))
p1 <- p1 + theme(axis.title=element_text(size=16))
p1 <- p1 + theme(plot.title=element_text(size=20, face="bold",hjust=0.5))
p1 <- p1 +  transition_time(year) + labs(title="Confirmed Planets as of {floor(frame_time)}")


animate(p1, nframes=300, width=1200, height=800, res=144, renderer=av_renderer())  #makes mp4         
#animate(p1, nframes=300, width=500, height=300, end_pause=30) #makes gif
```

```{r}
anim_save("plots/size_history.gif")
```


Doing this all again, separating kepler and non kepler...

```{r}
krad_planets <- rad_planets %>% filter(facility == "Kepler/K2 Discoveries")

temp <- krad_planets %>% filter(size == "Earth") 
earth_history <- count(temp, pl_disc)
colnames(earth_history) <- c("year", "Earths")
earth_history <- add_row(earth_history, year=1989:2008, Earths=0)  #adding in empty rows for gganimate

temp <- krad_planets %>% filter(size == "SuperEarth") 
sd_history <- count(temp, pl_disc)
colnames(sd_history) <- c("year", "SuperEarths")

temp <- krad_planets %>% filter(size == "Neptune") 
nep_history <- count(temp, pl_disc)
colnames(nep_history) <- c("year", "Neptunes")

temp <- krad_planets %>% filter(size == "Jupiter") 
jup_history <- count(temp, pl_disc)
colnames(jup_history) <- c("year","Jupiters")

temp <- krad_planets %>% filter(size == "Huge") 
huge_history <- count(temp, pl_disc)
colnames(huge_history) <- c("year", "HugePlanets")

disc_history <- count(krad_planets, pl_disc)
colnames(disc_history) <- c("year", "Planets")

ksize_history <- merge(disc_history, earth_history, by.x="year", by.y="year", all=TRUE)
ksize_history <- merge(ksize_history, sd_history, by.x="year",by.y="year",all.x=TRUE)
ksize_history <- merge(ksize_history, nep_history, by.x="year", by.y="year",all.x=TRUE)
ksize_history <- merge(ksize_history, jup_history, by.x="year", by.y="year", all.x=TRUE)
ksize_history <- merge(ksize_history, huge_history, by.x="year", by.y = "year", all.x=TRUE)

ksize_history$earth_cum <- cum_column(ksize_history$year, ksize_history$Earths)
ksize_history$sd_cum <- cum_column(ksize_history$year, ksize_history$SuperEarths)
ksize_history$nep_cum <- cum_column(ksize_history$year, ksize_history$Neptunes)
ksize_history$jup_cum <- cum_column(ksize_history$year, ksize_history$Jupiters)
ksize_history$huge_cum <- cum_column(ksize_history$year, ksize_history$HugePlanets)

temp <- ksize_history %>% select(c("year", "earth_cum", "sd_cum","nep_cum","jup_cum","huge_cum"))
colnames(temp) <- c("year","earth","superearth","neptune","jupiter","huge")

ksize_history_long <- gather(temp, size, count, earth:huge)


ksize_history_long$size_f <- factor(ksize_history_long$size, levels=c("earth","superearth","neptune","jupiter","huge"))

ksize_history_long$facility <- "Kepler/K2 Discoveries"

```


```{r}
nkrad_planets <- rad_planets %>% filter(facility == "All Other Discoveries")

temp <- nkrad_planets %>% filter(size == "Earth") 
earth_history <- count(temp, pl_disc)
colnames(earth_history) <- c("year", "Earths")
earth_history <- add_row(earth_history, year=1990:1991, Earths=0)
earth_history <- add_row(earth_history, year=1993, Earths=0)

temp <- nkrad_planets %>% filter(size == "SuperEarth") 
sd_history <- count(temp, pl_disc)
colnames(sd_history) <- c("year", "SuperEarths")

temp <- nkrad_planets %>% filter(size == "Neptune") 
nep_history <- count(temp, pl_disc)
colnames(nep_history) <- c("year", "Neptunes")

temp <- nkrad_planets %>% filter(size == "Jupiter") 
jup_history <- count(temp, pl_disc)
colnames(jup_history) <- c("year","Jupiters")

temp <- nkrad_planets %>% filter(size == "Huge") 
huge_history <- count(temp, pl_disc)
colnames(huge_history) <- c("year", "HugePlanets")

disc_history <- count(nkrad_planets, pl_disc)
colnames(disc_history) <- c("year", "Planets")

nksize_history <- merge(disc_history, earth_history, by.x="year", by.y="year", all=TRUE)
nksize_history <- merge(nksize_history, sd_history, by.x="year",by.y="year",all.x=TRUE)
nksize_history <- merge(nksize_history, nep_history, by.x="year", by.y="year",all.x=TRUE)
nksize_history <- merge(nksize_history, jup_history, by.x="year", by.y="year", all.x=TRUE)
nksize_history <- merge(nksize_history, huge_history, by.x="year", by.y = "year", all.x=TRUE)

nksize_history$earth_cum <- cum_column(nksize_history$year, nksize_history$Earths)
nksize_history$sd_cum <- cum_column(nksize_history$year, nksize_history$SuperEarths)
nksize_history$nep_cum <- cum_column(nksize_history$year, nksize_history$Neptunes)
nksize_history$jup_cum <- cum_column(nksize_history$year, nksize_history$Jupiters)
nksize_history$huge_cum <- cum_column(nksize_history$year, nksize_history$HugePlanets)

temp <- nksize_history %>% select(c("year", "earth_cum", "sd_cum","nep_cum","jup_cum","huge_cum"))
colnames(temp) <- c("year","earth","superearth","neptune","jupiter","huge")

nksize_history_long <- gather(temp, size, count, earth:huge)


nksize_history_long$size_f <- factor(nksize_history_long$size, levels=c("earth","superearth","neptune","jupiter","huge"))

nksize_history_long$facility <- "All Other Discoveries"

```


```{r}
comp_size_history_long <- rbind(ksize_history_long, nksize_history_long)
```

Will need to add a few lines for years with no data...

```{r}
#count(nksize_history_long, year)
```

```{r}
#comp_size_history_long <- add_row(comp_size_history_long, year=1990:1991, facility = "not Kepler",size="earth")
#comp_size_history_long <- add_row(comp_size_history_long, year=1993, facility = "not Kepler",size="earth")
```

```{r}
count(ksize_history_long, year)
```

```{r}
#comp_size_history_long <- add_row(comp_size_history_long, year=1989:2009, facility = "Kepler",size="earth")
```

Finally plot!
```{r}

comp_size_history_long$facility <- factor(comp_size_history_long$facility, 
                                          levels=c("Kepler/K2 Discoveries","All Other Discoveries"))

temp <- comp_size_history_long %>% filter(year == 2019)

baseplot <- ggplot(temp) + theme_bw()

p1 <- baseplot + geom_bar(aes(x=size_f, y=count, col=size_f,fill=size_f),stat="identity")
p1 <- p1 + scale_color_geert(palette="planets2") + scale_fill_geert(palette="planets2")
p1 <- p1 + xlab("size") + ylab("Number Known")
p1 <- p1 + theme(legend.position = "none")
p1 <- p1 + theme(axis.text=element_text(size=14))
p1 <- p1 + theme(axis.text.x=element_text(angle=90, hjust=1,vjust=0.5))
p1 <- p1 + theme(axis.title=element_text(size=16))
p1 <- p1 + theme(plot.title=element_text(size=20, face="bold",hjust=0.5))
p1 <- p1 + theme(strip.text.x = element_text(size = 14))
p1 <- p1 + facet_grid(aes(cols=facility))
print(p1)
```


And animate...


```{r}
baseplot <- ggplot(comp_size_history_long) + theme_bw()

p1 <- baseplot + geom_bar(aes(x=size_f, y=count, col=size_f,fill=size_f),stat="identity")
p1 <- p1 + scale_color_geert(palette="planets2") + scale_fill_geert(palette="planets2")
p1<- p1 + xlab("size") + ylab("Number Known")
p1 <- p1 + theme(legend.position = "none")
p1 <- p1 + theme(axis.text=element_text(size=14))
p1 <- p1 + theme(axis.text.x=element_text(angle=90, hjust=1,vjust=0.5))
p1 <- p1 + theme(axis.title=element_text(size=16))
p1 <- p1 + theme(plot.title=element_text(size=20, face="bold",hjust=0.5))
p1 <- p1 + theme(strip.text.x = element_text(size = 14))
p1 <- p1 + facet_grid(aes(cols=facility))
p1 <- p1 +  transition_time(year) + 
  labs(title="Confirmed Planets as of {floor(frame_time)}")


animate(p1, nframes=300, width=1200, height=800, res=144, renderer=av_renderer())  #makes mp4         
#animate(p1, nframes=300, width=800, height=500, end_pause=30) #makes gif
```

```{r}
anim_save("plots/comp_size_history.gif")
```





