---
title: "Animated Planets"
author: "J. Dotson"
date: "6/15/2019"
output: html_document
---

Animated plots showing evolution of confirmed planets over time...


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

require(ggplot2)
require(dplyr)
require(tidyverse)
require(readr)
require(RColorBrewer)
require(ggforce)
require(ggthemes)
require(gridExtra)
require(gganimate)
require(tidyr)

myecho=FALSE
knitr::opts_knit$set(root.dir = '../.')
```

Using the planet property table created in planethistory_base.  

```{r }

p_planets <- readRDS("RDS/planets_to_plot.rds")

```

Starting with our usual scatter plot...

```{r }

baseplot<-ggplot(p_planets) + theme_bw() +  ylab("Size Relative to Earth (Radius)") +
  xlab("Orbital Period (Days)")

myylim<-c(0.3,25)
myxlim<-c(0.1,1000)

baseplot2 <- baseplot + scale_x_log10(limits=myxlim,breaks=c(1,10,100,1000)) + 
  scale_y_log10(limits=myylim,breaks=c(1,4,10,20,40))

thisplot <- baseplot2 + geom_point(aes(x=pl_orbper,y=new_rade,group=seq_along(pl_disc)),
                                   shape=21,fill="purple3",col="white",size=4)
#
print(thisplot)

thisplot <- thisplot + transition_time(pl_disc) + labs(title="Year: {floor(frame_time)}")
thisplot <- thisplot + shadow_mark(past=TRUE, col="white",fill="darkgrey") + enter_appear()
thisplot <- thisplot + theme(plot.title=element_text(size=16),
                             axis.title.x=element_text(size=14),
                             axis.title.y=element_text(size=14))


# this call to animate will make an mp4
#animate(thisplot, nframes = 300,renderer = av_renderer(),width=1200, height=800, res=144)

# this call to animate will make a gif
animate(thisplot, nframes=300, width=1200, height=800, res=144) 




```


```{r}
anim_save("plots/scatter_history.gif")
```


Let's sort the planets into size bins...

Using the bins we've used in our public materials (btw, NExScI uses these bins also...)

Earth-size:  < 1.25 Re
Super earth size: 1.25 - 2Re
Neptune size: 2 - 6 Re
Jupiter size: 6 - 15 Re
larger: 15 - 25 Re


mutate(gradebook, letter = ifelse(grade %in% 60:69, "D",
                                     ifelse(grade %in% 70:79, "C",
                                            ifelse(grade %in% 80:89, "B",
                                                   ifelse(grade %in% 90:99, "A", "F")))))

```{r}

earth_max <- 1.25
super_earth_max <- 2.0
neptune_max <- 6.0
jupiter_max <- 15.0

p_planets$size <- rep(NA, nrow(p_planets))

p_planets <- mutate(p_planets, size = ifelse(((new_rade > neptune_max) & (new_rade <= jupiter_max)), "Jupiter",
                                             ifelse(((new_rade > super_earth_max) & (new_rade <= neptune_max)), "Neptune",
                                                    ifelse(((new_rade > earth_max)& (new_rade <= super_earth_max)), "SuperEarth",
                                                           ifelse(((new_rade > 0) & (new_rade <= earth_max)), "Earth", "Huge")))))


```




```{r}

#just for now, remove the planets without a radius...
rad_planets <- p_planets %>% filter(!is.na(size))

rad_planets$size <- factor(rad_planets$size, levels=c("Earth","SuperEarth","Neptune","Jupiter","Huge"))

baseplot <- ggplot(rad_planets) + theme_bw()

bar_p <- baseplot + geom_bar(aes(size,col=size, fill=size))

bar_p<- bar_p + xlab("") + ylab("Number Known")
bar_p <- bar_p + theme(legend.position = "none")
bar_p <- bar_p + theme(axis.text=element_text(size=14))
bar_p <- bar_p + theme(axis.title=element_text(size=16))
bar_p <- bar_p + theme(plot.title=element_text(size=20, face="bold",hjust=0.5))


print(bar_p)
```



To animate, we need to reorganize the data bit so that the cumulative counts per year are available. 

Need column for count, size, and year.  

First figure out how many of each size were discovered in a given year.


```{r}
temp <- rad_planets %>% filter(size == "Earth")
earth_history <- count(temp, pl_disc)
colnames(earth_history) <- c("year", "Earths")

temp <- rad_planets %>% filter(size == "SuperEarth")
sd_history <- count(temp, pl_disc)
colnames(sd_history) <- c("year", "SuperEarths")

temp <- rad_planets %>% filter(size == "Neptune")
nep_history <- count(temp, pl_disc)
colnames(nep_history) <- c("year", "Neptunes")

temp <- rad_planets %>% filter(size == "Jupiter")
jup_history <- count(temp, pl_disc)
colnames(jup_history) <- c("year","Jupiters")

temp <- rad_planets %>% filter(size == "Huge")
huge_history <- count(temp, pl_disc)
colnames(huge_history) <- c("year", "HugePlanets")

disc_history <- count(rad_planets, pl_disc)
colnames(disc_history) <- c("year", "Planets")

size_history <- merge(disc_history, earth_history, by.x="year", by.y="year", all.x=TRUE)
size_history <- merge(size_history, sd_history, by.x="year",by.y="year",all.x=TRUE)
size_history <- merge(size_history, nep_history, by.x="year", by.y="year",all.x=TRUE)
size_history <- merge(size_history, jup_history, by.x="year", by.y="year", all.x=TRUE)
size_history <- merge(size_history, huge_history, by.x="year", by.y = "year", all.x=TRUE)




```

Next calculate the cumulative number in each size bin as a function of time.

```{r}
cum_column <- function(myx, myy){
  myresult <- tibble(x = myx, y=myy)
  jj <- which(is.na(myresult$y))
  myresult$y[jj] <- 0
  myresult$cum <- rep(0,nrow(myresult))
  myresult$cum[1] <- myresult$y[1]
  for (i in 2:length(myx)){
    myresult$cum[i] <- myresult$cum[i-1] + myresult$y[i]
  }
  return(myresult$cum)
}
```


```{r}
size_history$earth_cum <- cum_column(size_history$year, size_history$Earths)
size_history$sd_cum <- cum_column(size_history$year, size_history$SuperEarths)
size_history$nep_cum <- cum_column(size_history$year, size_history$Neptunes)
size_history$jup_cum <- cum_column(size_history$year, size_history$Jupiters)
size_history$huge_cum <- cum_column(size_history$year, size_history$HugePlanets)
```


And now put the cumulative counts into a tidy long format
```{r}

temp <- size_history %>% select(c("year", "earth_cum", "sd_cum","nep_cum","jup_cum","huge_cum"))
colnames(temp) <- c("year","earths","superearths","neptunes","jupiters","larger")

size_history_long <- gather(temp, size, count, earths:larger)


size_history_long$size_f <- factor(size_history_long$size, levels=c("earths","superearths","neptunes","jupiters","larger"))

```


Finally plot!
```{r}

temp <- size_history_long %>% filter(year == 2019)

baseplot <- ggplot(temp) + theme_bw()

p1 <- baseplot + geom_bar(aes(x=size_f, y=count, col=size,fill=size),stat="identity")
p1<- p1 + xlab("") + ylab("Number Known")
p1 <- p1 + theme(legend.position = "none")
p1 <- p1 + theme(axis.text=element_text(size=14))
p1 <- p1 + theme(axis.title=element_text(size=16))
p1 <- p1 + theme(plot.title=element_text(size=20, face="bold",hjust=0.5))
print(p1)
```

And animate...


```{r}
baseplot <- ggplot(size_history_long) + theme_bw()

p1 <- baseplot + geom_bar(aes(x=size_f, y=count, col=size,fill=size),stat="identity")
p1<- p1 + xlab("") + ylab("Number Known")
p1 <- p1 + theme(legend.position = "none")
p1 <- p1 + theme(axis.text=element_text(size=14))
p1 <- p1 + theme(axis.title=element_text(size=16))
p1 <- p1 + theme(plot.title=element_text(size=20, face="bold",hjust=0.5))
p1 <- p1 +  transition_time(year) + labs(title="Year: {floor(frame_time)}")


#animate(p1, nframes=300, width=1200, height=800, res=144, renderer=av_renderer())  #makes mp4         
animate(p1, nframes=300, width=600, height=500, end_pause=30) #makes gif
```

```{r}
anim_save("plots/size_history.gif")
```



