---
title: "planet history"
author: "J. Dotson"
date: "12/3/2017"
output: html_document
---

Updated May 2019.  No longer need to manually add in Kepler-90 and Kepler-80 discoveries.  Purpose here is to put together end of emission plots.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(ggplot2)
require(dplyr)
require(tidyverse)
require(readr)
require(RColorBrewer)
require(ggforce)
require(ggthemes)
require(gridExtra)
myecho=FALSE
knitr::opts_knit$set(root.dir = '../.')

```

```{r, echo=myecho}
source("Rscripts/mythemes.R")
```



## K2 Exoplanet stats

the point of this document is to set up a script I can use over (and over) to generate some K2 stats and make some plots.


Start by downloading the confirmed planet table from NExScI.

```{r download, echo=myecho}
baseurl<-"https://exoplanetarchive.ipac.caltech.edu/cgi-bin/nstedAPI/nph-nstedAPI?table=exoplanets"
params<-"pl_hostname,pl_letter,pl_discmethod,pl_pnum,pl_orbper,pl_orbsmax,pl_orbsmaxerr1,pl_orbsmaxerr2,pl_orbeccen,pl_kepflag,pl_k2flag,ra_str,dec_str,ra,dec,st_dist,st_optmag,st_optband,gaia_gmag,st_teff,st_tefferr1,st_tefferr2,st_mass,st_masserr1,st_masserr2,st_rad,st_raderr1,st_raderr2,pl_cbflag,pl_eqt,pl_insol,pl_bmasse,pl_bmasseerr1,pl_bmasseerr2,pl_msinie,pl_rade,pl_radeerr1,pl_radeerr2,pl_trandep,pl_trandur,pl_imppar,pl_disc,pl_status,gaia_dist,st_spstr,st_logg,st_metfe,st_age,pl_locale,pl_facility,pl_telescope,pl_instrument,rowupdate"


address<-paste(baseurl,"&select=",params,sep="")

download.file(address, "data/confirmedplanets.csv")

print(paste("Confirmed planets table downloaded on ",date()))
```


```{r inputs, echo=myecho}
# when there's new data, need to update filename and date
confirmedplanets <- read_csv("data/confirmedplanets.csv", comment = "#")

confirmedplanets$pl_masse <- confirmedplanets$pl_bmasse  #check that this is ok!


###Ugh, going back to manually downloading the table for expediency sake :(  6/21/18

#confirmedplanets <- read_csv("inputs/planets180621.csv", comment = "#")

```



```{r ,echo=myecho}
mytheme<-theme_bw()
#mytheme2 <- theme_hc(bgcolor="darkunica")
#mytheme3 <- theme_solarized(light=FALSE)

plothist<- ggplot(confirmedplanets) + geom_bar(aes(pl_disc)) 
#print(plothist + mytheme)
#print(plothist + mytheme2)
#print(plothist + mytheme3)
print(plothist + theme_kepler())
#print(plothist + theme_ack())
ggsave("plots/test.eps")

```

## Need to estimate radii for RV planets

Ideally i'd like to use Chen & Kipling's forecaster, but since this is just for show, let's simplify and not stress about how to go back and forth between python and R....

1) pull a random inclination between 75 and 90 to get the true Mass.  (why this range?)

2) choose which power law to use from chen & kipping
 < 2.0 Me --> Terran
 > Terran & < 0.41 Mj --> Neptunian
 > Neptunian & < 0.08 Msol --> Jovian

3) calculate radius from appropriate power law
  Terran:  R = 1.008 * M^0.28
  Neptunian: R = ?? * M^0.59
  Jovian:  R = ?? * M^-0.04
  
4) add the appropriate fractional noise.  Namely, rechoose a radius from a distribution centered at the radius calculated above and a sigma given by:
  Terran: sigma = 0.04 * R
  Neptunian:  sigma = 0.146 * R
  Jovian: sigma = 0.0737 * R
 

```{r, echo=myecho}
#first testing on a synthetic population...
solarsystem<-read_csv("inputs/solarsystem.csv", comment = "#")
Mass_earth <- 5.97e24 #kg
Rad_earth <- 12756/2 #km
Mfake<-seq(0.5, 500, by=0.5)

Ct <- 1.008
St <- 0.28
sigmaT <- 0.04
Mmax_t <- 2.0
Sn <- 0.59
Cn <- Ct * Mmax_t^St / Mmax_t^Sn  #lines have to meet
sigmaN <- 0.146
Mmax_n <- 0.41 * 317.83 
Sj <- -0.04
Cj <- Cn * Mmax_n^Sn / Mmax_n^Sj
sigmaJ <- 0.0737
sigmaJ <- 2*sigmaJ  # making it bigger, because...

Terran_radii <- Ct * Mfake^St

fakeplanets<-data.frame(mass=Mfake,terran=Terran_radii)

fakeplanets$neptunian <- Cn * Mfake^Sn
fakeplanets$jovian <- Cj * Mfake^Sj

p1<-ggplot(fakeplanets) + geom_point(aes(x=mass,y=terran),col="blue") + scale_x_log10() + scale_y_log10()
p1 <- p1 + geom_point(aes(x=mass,y=neptunian),col="green")
p1 <- p1 + geom_point(aes(x=mass,y=jovian),col="red")
print(p1)

```

```{r, echo=myecho}
rvpl<-filter(confirmedplanets,pl_discmethod == "Radial Velocity")
my_is <- 0
for (i in 1: nrow(rvpl)){
  if(is.na(rvpl$pl_masse[i])){
    this_i <- runif(1,min=15,max=90)
    my_is <- c(my_is,this_i)
    rvpl$pl_masse[i] <- rvpl$pl_msinie[i] / sin(this_i/360 * 2* 3.14159)
  }
}

my_is <- my_is[-1]
#hist(my_is)

rvpl <- filter(rvpl, !is.na(pl_masse))

rvpl$pl_Rforecast <- rep(0,nrow(rvpl))
rvpl$pl_Rforecast2 <- rep(0,nrow(rvpl))

for (i in 1:nrow(rvpl)){
  if(rvpl$pl_masse[i] <= Mmax_t) {
    rvpl$pl_Rforecast[i] <- Ct * rvpl$pl_masse[i]^St
    thissig <- sigmaT * rvpl$pl_Rforecast[i]
    rvpl$pl_Rforecast2[i] <- rnorm(1,mean=rvpl$pl_Rforecast[i],sd=thissig)
  } else if (rvpl$pl_masse[i] <= Mmax_n){
    rvpl$pl_Rforecast[i] <- Cn * rvpl$pl_masse[i]^Sn
    thissig <- sigmaN * rvpl$pl_Rforecast[i]
    rvpl$pl_Rforecast2[i] <- rnorm(1,mean=rvpl$pl_Rforecast[i],sd=thissig)    
  } else {
    rvpl$pl_Rforecast[i] <- Cj * rvpl$pl_masse[i]^Sj
    thissig <- sigmaJ * rvpl$pl_Rforecast[i]
    rvpl$pl_Rforecast2[i] <- rnorm(1,mean=rvpl$pl_Rforecast[i],sd=thissig)
  }
}


```

```{r, echo=myecho}

p2 <- ggplot(rvpl) + geom_point(aes(x=pl_masse,y=pl_Rforecast)) + 
  geom_point(aes(x=pl_masse,y=pl_Rforecast2),col="cadetblue")
p2 <- p2 + scale_x_log10() + scale_y_log10()
print(p2)


```

Let's check and see how many missing values we now have...

```{r}
temp <- confirmedplanets %>% filter(is.na(pl_rade))
count(temp,pl_discmethod)
temp2 <- temp %>% filter(pl_discmethod != "Radial Velocity")

temp3 <- rvpl %>% filter(is.na(pl_Rforecast))

temp4 <- confirmedplanets %>% filter(is.na(pl_orbper))
count(temp4,pl_discmethod)
```

So, the planets without orbital period are, given their discovery method just fine.  And it also looks like I manage to successfully create a radii for all the RV planets.  BUT, I'm not happy with all of the planets missing radii with other detection mechanisms.  Back down at the bottom, I'm going to look at filling in some of those from the composite table.  (Which has a lot fewer)




```{r, echo=myecho}


prekepler<-filter(confirmedplanets, pl_disc < 2009)
prekepler_rv <- filter(rvpl,pl_disc < 2009)


baseplot<-ggplot(prekepler) + theme_kepler() + ylab("Size Relative to Earth (Radius)") +
  xlab("Orbital Period (Days)")

myylim<-c(0.3,25)
myxlim<-c(0.1,1000)
mycol1<-"blue"
mycol2<-"yellow"
mycol5<-"pink"
mycol3<-mycol2
mycol4<-mycol1
baseplot2 <- baseplot + scale_x_log10(limits=myxlim,breaks=c(1,10,100,1000)) + 
  scale_y_log10(limits=myylim,breaks=c(1,4,10,20,40))
baseplot3 <- baseplot2 + geom_segment(aes(x=0.1,xend=1000,y=1,yend=1),col="white",size=0.1)
baseplot3 <- baseplot3 + geom_segment(aes(x=0.1,xend=1000,y=3.86, yend=3.86),col="white",size=0.1)
baseplot3 <- baseplot3 + geom_segment(aes(x=0.1,xend=1000,y=10.97,yend=10.97),col="white",size=0.1)

preplot <- baseplot3 + geom_point(aes(x=pl_orbper,y=pl_rade),shape=21,fill=mycol1,col="black",size=2)
preplot <- preplot + geom_point(data=prekepler_rv,aes(x=pl_orbper,y=pl_Rforecast2),
                                shape=21,fill=mycol4,col="black",size=2)
print(preplot)
ggsave("plots/pre_kepler.eps")

postkepler<-filter(confirmedplanets, pl_disc >=2009)
postkepler_rv <-filter(rvpl,pl_disc >= 2009)

postplot <- preplot + geom_point(data=postkepler_rv,aes(x=pl_orbper,y=pl_Rforecast2),
                                  shape=21,fill=mycol3,col="black", size=2)
postplot <- postplot + geom_point(data=postkepler,aes(x=pl_orbper,y=pl_rade),
                                 shape=21,fill=mycol2,col="black",size=2)

print(postplot)
ggsave("plots/post_kepler.eps")

kepler<-filter(confirmedplanets,pl_kepflag == 1)
k2 <- filter(confirmedplanets, pl_k2flag ==1)

postplot2 <- preplot + geom_point(data=kepler,aes(x=pl_orbper,y=pl_rade),
                                 shape=21,fill=mycol2,col="black",size=2)
postplot2 <- postplot2 + geom_point(data=k2,aes(x=pl_orbper,y=pl_rade),
                                 shape=21,fill=mycol2,col="black",size=2)

print(postplot2)
ggsave("plots/kepleronly.eps")


notkep_post <- filter(postkepler,pl_kepflag == 0)
notkep_post <- filter(notkep_post,pl_k2flag == 0)
notkeplerplot<- preplot + geom_point(data=notkep_post,aes(x=pl_orbper,y=pl_rade),
                                 shape=21,fill=mycol5,col="black",size=2) 
notkeplerplot <- notkeplerplot + geom_point(data=postkepler_rv,aes(x=pl_orbper,y=pl_Rforecast2),
                                  shape=21,fill=mycol5,col="black", size=2)
print(notkeplerplot)
ggsave("plots/notkepler.eps")

pastaplot<- notkeplerplot + geom_point(data=kepler,aes(x=pl_orbper,y=pl_rade),
                                 shape=21,fill=mycol2,col="black",size=2)
pastaplot <- pastaplot + geom_point(data=k2,aes(x=pl_orbper,y=pl_rade),
                                 shape=21,fill=mycol2,col="black",size=2)

print(pastaplot)
ggsave("plots/post_kepler2.eps")

```

July 2018, redo above more generally.  Also -- want to find out why there are missing rows!

```{r}

prekepler<-filter(confirmedplanets, pl_disc < 2009)
count(prekepler, is.na(pl_orbper))
jj <- is.na(prekepler$pl_orbper)
prekep_missing_orbper <- prekepler[jj,]
count(prekep_missing_orbper, pl_discmethod)
print("The planets missing orbital period are all either Imaging or microlensing.  Think we have to let those go right now...")

count(prekepler, is.na(pl_rade))
jj <- is.na(prekepler$pl_rade)
prekep_missing_rade <- prekepler[jj,]
count(prekep_missing_rade,pl_discmethod)
print("These are mostly radial velocity planets.  So, only need to worry about the ones that aren't in the RV dataframe.")
notmatched <- anti_join(prekepler,prekepler_rv,by="pl_hostname")
count(notmatched,pl_discmethod)


prekepler_rv <- filter(rvpl,pl_disc < 2009)
count(prekepler_rv, is.na(pl_orbper))
count(prekepler_rv, is.na(pl_Rforecast2))


```





```{r, echo=myecho}

prekepler<-filter(confirmedplanets, pl_disc < 2009)
prekepler_rv <- filter(rvpl,pl_disc < 2009)


baseplot<-ggplot(prekepler) + theme_kepler() + ylab("Size Relative to Earth (Radius)") +
  xlab("Orbital Period (Days)")

myylim<-c(0.3,25)
myxlim<-c(0.1,1000)
mycol1<-"blue"
mycol2<-"yellow"
mycol5<-"pink"
mycol3<-mycol2
mycol4<-mycol1
baseplot2 <- baseplot + scale_x_log10(limits=myxlim,breaks=c(1,10,100,1000)) + 
  scale_y_log10(limits=myylim,breaks=c(1,4,10,20,40))
baseplot3 <- baseplot2 + geom_segment(aes(x=0.1,xend=1000,y=1,yend=1),col="white",size=0.1)
baseplot3 <- baseplot3 + geom_segment(aes(x=0.1,xend=1000,y=3.86, yend=3.86),col="white",size=0.1)
baseplot3 <- baseplot3 + geom_segment(aes(x=0.1,xend=1000,y=10.97,yend=10.97),col="white",size=0.1)

preplot <- baseplot3 + geom_point(aes(x=pl_orbper,y=pl_rade),shape=21,fill=mycol1,col="black",size=2)
preplot <- preplot + geom_point(data=prekepler_rv,aes(x=pl_orbper,y=pl_Rforecast2),
                                shape=21,fill=mycol4,col="black",size=2)
print(preplot)
ggsave("plots/pre_kepler.eps")





postkepler<-filter(confirmedplanets, pl_disc >=2009)
postkepler_rv <-filter(rvpl,pl_disc >= 2009)

postplot <- preplot + geom_point(data=postkepler_rv,aes(x=pl_orbper,y=pl_Rforecast2),
                                  shape=21,fill=mycol3,col="black", size=2)
postplot <- postplot + geom_point(data=postkepler,aes(x=pl_orbper,y=pl_rade),
                                 shape=21,fill=mycol2,col="black",size=2)


print(postplot)
ggsave("plots/post_kepler.eps")

kepler<-filter(confirmedplanets,pl_kepflag == 1)
k2 <- filter(confirmedplanets, pl_k2flag ==1)

postplot2 <- preplot + geom_point(data=kepler,aes(x=pl_orbper,y=pl_rade),
                                 shape=21,fill=mycol2,col="black",size=2)
postplot2 <- postplot2 + geom_point(data=k2,aes(x=pl_orbper,y=pl_rade),
                                 shape=21,fill=mycol2,col="black",size=2)

print(postplot2)
ggsave("plots/kepleronly.eps")


notkep_post <- filter(postkepler,pl_kepflag == 0)
notkep_post <- filter(notkep_post,pl_k2flag == 0)
notkeplerplot<- preplot + geom_point(data=notkep_post,aes(x=pl_orbper,y=pl_rade),
                                 shape=21,fill=mycol5,col="black",size=2) 
notkeplerplot <- notkeplerplot + geom_point(data=postkepler_rv,aes(x=pl_orbper,y=pl_Rforecast2),
                                  shape=21,fill=mycol5,col="black", size=2)
print(notkeplerplot)
ggsave("plots/notkepler.eps")

pastaplot<- notkeplerplot + geom_point(data=kepler,aes(x=pl_orbper,y=pl_rade),
                                 shape=21,fill=mycol2,col="black",size=2)
pastaplot <- pastaplot + geom_point(data=k2,aes(x=pl_orbper,y=pl_rade),
                                 shape=21,fill=mycol2,col="black",size=2)

print(pastaplot)
ggsave("plots/post_kepler2.eps")

```





```{r}
prekepler<-filter(confirmedplanets, pl_disc < 2009)
prekepler_rv <- filter(rvpl,pl_disc < 2009)


baseplot<-ggplot(prekepler) + theme_bw() + ylab("Size Relative to Earth (Radius)") +
  xlab("Orbital Period (Days)")

myylim<-c(0.3,25)
myxlim<-c(0.1,1000)
mycol1<-"blue"
mycol2<-"yellow"
mycol5<-"pink"
mycol3<-mycol2
mycol4<-mycol1
baseplot2 <- baseplot + scale_x_log10(limits=myxlim,breaks=c(1,10,100,1000)) + 
  scale_y_log10(limits=myylim,breaks=c(1,4,10,20,40))
#baseplot3 <- baseplot2 + geom_segment(aes(x=0.1,xend=1000,y=1,yend=1),col="white",size=0.1)
#baseplot3 <- baseplot3 + geom_segment(aes(x=0.1,xend=1000,y=3.86, yend=3.86),col="white",size=0.1)
#baseplot3 <- baseplot3 + geom_segment(aes(x=0.1,xend=1000,y=10.97,yend=10.97),col="white",size=0.1)

preplot <- baseplot2 + geom_point(aes(x=pl_orbper,y=pl_rade),shape=21,size=2,fill="cadetblue")
preplot <- preplot + geom_point(data=prekepler_rv,aes(x=pl_orbper,y=pl_Rforecast2),
                                shape=21,size=2,fill="grey75")
print(preplot)
#ggsave("plots/pre_kepler.eps")

postkepler<-filter(confirmedplanets, pl_disc >=2009)
postkepler_other <- filter(postkepler, pl_instrument != "Kepler CCD Array")
kepler <- filter(confirmedplanets, pl_instrument == "Kepler CCD Array")
postkepler_rv <-filter(rvpl,pl_disc >= 2009)

postplot <- preplot + geom_point(data=postkepler_rv,aes(x=pl_orbper,y=pl_Rforecast2),
                                  shape=21,fill="mediumpurple", size=2)
postplot <- postplot + geom_point(data=postkepler_other, aes(x=pl_orbper,y=pl_rade),
                                  shape=21,fill="mediumpurple", size=2)
postplot <- postplot + geom_point(data=kepler,aes(x=pl_orbper,y=pl_rade),
                                 shape=21,fill="magenta",size=2)


print(postplot)
ggsave("plots/post_kepler.jpg")

postkepler<-filter(confirmedplanets, pl_disc >=2009)
postkepler_other <- filter(postkepler, pl_instrument != "Kepler CCD Array")
kepler <- filter(confirmedplanets, pl_facility == "Kepler")
k2 <- filter(confirmedplanets, pl_facility == "K2")
postkepler_rv <-filter(rvpl,pl_disc >= 2009)

postplot <- preplot + geom_point(data=postkepler_rv,aes(x=pl_orbper,y=pl_Rforecast2),
                                  shape=21,fill="mediumpurple", size=2)
postplot <- postplot + geom_point(data=postkepler_other, aes(x=pl_orbper,y=pl_rade),
                                  shape=21,fill="mediumpurple", size=2)
postplot <- postplot + geom_point(data=kepler,aes(x=pl_orbper,y=pl_rade),
                                 shape=21,fill="magenta",size=2)

postplot <- postplot + geom_point(data=k2,aes(x=pl_orbper,y=pl_rade),
                                 shape=21,fill="turquoise1",size=2)



print(postplot)
ggsave("plots/forsteve.jpg")




```


So the above is all well and good, but I'm thinking the correct way to do this *might* be to use the NExScI composite table rather than roll my own radius / masses.  Let's just check that out a wee bit...  (Looks like I am going to need to join with the confirmed table in order to get the kepler/k2 flags...)


```{r}

baseurl<-"https://exoplanetarchive.ipac.caltech.edu/cgi-bin/nstedAPI/nph-nstedAPI?table=compositepars"
params<-"fpl_hostname,fpl_letter,fpl_name,fpl_discmethod,fpl_disc,fpl_orbper,fpl_smax,fpl_eccen,fpl_bmasse,fpl_bmassreflink,fpl_rade,fpl_radreflink"

#I'd like to pull the references down, but there are some bad characters which munge up my ability to read in things afterwards.  Let's try just putting the references at the end and see if that fixes things...
params<-"fpl_hostname,fpl_letter,fpl_name,fpl_discmethod,fpl_disc,fpl_orbper,fpl_smax,fpl_eccen,fpl_bmasse,fpl_rade,fpl_bmassreflink,fpl_radreflink"

address<-paste(baseurl,"&select=",params,sep="")

download.file(address, "data/compositeplanets.csv")

print(paste("Composite planets table downloaded on ",date()))

```

```{r}
compositeplanets <- read_csv("data/compositeplanets.csv", comment = "#")

compositeplanets$my_name <- paste(compositeplanets$fpl_hostname, compositeplanets$fpl_letter)
```

```{r}
confirmedplanets$pl_name <- paste(confirmedplanets$pl_hostname, confirmedplanets$pl_letter)

confirmed_simplified <- confirmedplanets %>% select(pl_hostname,pl_letter,pl_name,pl_orbper,pl_facility)
```

Learned (again!) that the name field in the composite planets table is not always the same as the hostname + planet letter.  Ugh!


```{r}
planets <- merge(compositeplanets, confirmed_simplified, by.x="my_name",by.y="pl_name",all=TRUE)
```


Ok, now let's see if we're still missing sizes and periods

```{r}
# things without an orbital period are imaging, microlensing and pulsar timing.
jj <- is.na(planets$fpl_orbper)
temp <- planets[jj,]
count(temp, fpl_discmethod)

# there are a small number without radii.  The reasons aren't as obvious -- but a spot check tells me to move on
# for instance, even though wasp-53c is a transiting planet, neither the composite nor confirmed
# planet tables include a radius...  
jj <- is.na(planets$fpl_rade)
temp <- planets[jj,]
count(temp,fpl_discmethod)

```

We're losing a modest number of planets due to missing values in the composite table, but lets just go with it...


I tend to say prekepler is anything before 2009, but kepler planets didn't start getting published until 2010.  Just out of curiousity -- how many planets were discovered in 2009?



```{r}
temp <- filter(planets, fpl_disc == 2009)
print(paste(nrow(temp), "planets were discovered in 2009"))

temp <- filter(planets, fpl_disc < 2009)
print(paste(nrow(temp), "planets were discovered before 2009"))

temp <- filter(planets,fpl_disc <= 2009)
```

Nope -- I was wrong!!!  The first three Kepler planets are labeled as discovered in 2009 by NExScI.  Let's stick with the prior to launch (modulo 2 months) count.


```{r}
prekepler<-filter(planets, fpl_disc < 2009)



baseplot<-ggplot(prekepler) + theme_kepler() + ylab("Size Relative to Earth (Radius)") +
  xlab("Orbital Period (Days)")

myylim<-c(0.3,25)
myxlim<-c(0.1,1000)
mycol1<-"blue"
mycol2<-"yellow"
mycol5<-"pink"
mycol3<-mycol2
mycol4<-mycol1
baseplot2 <- baseplot + scale_x_log10(limits=myxlim,breaks=c(1,10,100,1000)) + 
  scale_y_log10(limits=myylim,breaks=c(1,4,10,20,40))
baseplot3 <- baseplot2 + geom_segment(aes(x=0.1,xend=1000,y=1,yend=1),col="white",size=0.1)
baseplot3 <- baseplot3 + geom_segment(aes(x=0.1,xend=1000,y=3.86, yend=3.86),col="white",size=0.1)
baseplot3 <- baseplot3 + geom_segment(aes(x=0.1,xend=1000,y=10.97,yend=10.97),col="white",size=0.1)

preplot <- baseplot3 + geom_point(aes(x=fpl_orbper,y=fpl_rade),shape=21,fill=mycol1,col="black",size=2)

print(preplot)
ggsave("plots/pre_kepler.eps")





postkepler<-filter(planets, fpl_disc >=2009)

postplot <- preplot + geom_point(data=postkepler,aes(x=fpl_orbper,y=fpl_rade),
                                 shape=21,fill=mycol2,col="black",size=2)
print(postplot)
ggsave("plots/post_kepler.eps")


kepler <- filter(planets,pl_facility == "Kepler")

postplot2 <- preplot + geom_point(data=kepler,aes(x=fpl_orbper,y=fpl_rade),
                                 shape=21,fill=mycol2,col="black",size=2)


print(postplot2)
ggsave("plots/kepleronly.eps")


notkep_post <- filter(postkepler,pl_facility != "Kepler")

notkeplerplot<- preplot + geom_point(data=notkep_post,aes(x=fpl_orbper,y=fpl_rade),
                                 shape=21,fill=mycol5,col="black",size=2) 

print(notkeplerplot)
ggsave("plots/notkepler.eps")

pastaplot<- notkeplerplot + geom_point(data=kepler,aes(x=fpl_orbper,y=fpl_rade),
                                 shape=21,fill=mycol2,col="black",size=2)


print(pastaplot)
ggsave("plots/post_kepler2.eps")

```


So, from a presentation perspective, I don't like how the NExScI derived radii are so tight.  But if I just use the confirmed planet table and my predictions, I'm missing the sizes of a bunch of non-rv planets.  Let's make a composite-composite table...

```{r}

#I want fewer columns so I don't get confused...
simpleplanets <- confirmedplanets %>% select(pl_hostname,pl_letter,pl_discmethod,pl_orbper,pl_rade,pl_disc,pl_facility,pl_name)

#First let's separate out the non-rv planets from the confirmed
non_rv <- simpleplanets %>% filter(pl_discmethod != "Radial Velocity")

# and identify the subset of those which do, indeed have pl_rade
non_rv_withsize <- non_rv %>% filter(!is.na(pl_rade))

# and now the non-rv planets without pl_rade
missingsize <- non_rv %>% filter(is.na(pl_rade))

# merge with composite
addsize <- merge(missingsize,compositeplanets, by.x="pl_name",by.y="my_name")
# replace rade from confirmed with rade from compsite
addsize$pl_rade <- addsize$fpl_rade
# select the simpleplanet columns
newlysized <- addsize %>% select(pl_hostname,pl_letter,pl_discmethod,pl_orbper,pl_rade,pl_disc,pl_facility,pl_name)

#put these back together
frankenplanets <- rbind(non_rv_withsize,newlysized)




```


Ok, now let's plot up like we used to...

```{r}
prekepler<-filter(frankenplanets, pl_disc < 2009)
prekepler_rv <- filter(rvpl,pl_disc < 2009)


baseplot<-ggplot(prekepler) + theme_kepler() + ylab("Size Relative to Earth (Radius)") +
  xlab("Orbital Period (Days)")

myylim<-c(0.3,25)
myxlim<-c(0.1,1000)
mycol1<-"blue"
mycol2<-"yellow"
mycol5<-"pink"
mycol3<-mycol2
mycol4<-mycol1
baseplot2 <- baseplot + scale_x_log10(limits=myxlim,breaks=c(1,10,100,1000)) + 
  scale_y_log10(limits=myylim,breaks=c(1,4,10,20,40))
baseplot3 <- baseplot2 + geom_segment(aes(x=0.1,xend=1000,y=1,yend=1),col="white",size=0.1)
baseplot3 <- baseplot3 + geom_segment(aes(x=0.1,xend=1000,y=3.86, yend=3.86),col="white",size=0.1)
baseplot3 <- baseplot3 + geom_segment(aes(x=0.1,xend=1000,y=10.97,yend=10.97),col="white",size=0.1)

preplot <- baseplot3 + geom_point(aes(x=pl_orbper,y=pl_rade),shape=21,fill=mycol1,col="black",size=2)
preplot <- preplot + geom_point(data=prekepler_rv,aes(x=pl_orbper,y=pl_Rforecast2),
                                shape=21,fill=mycol4,col="black",size=2)
print(preplot)
ggsave("plots/pre_kepler.eps")





postkepler<-filter(frankenplanets, pl_disc >=2009)
postkepler_rv <-filter(rvpl,pl_disc >= 2009)

postplot <- preplot + geom_point(data=postkepler_rv,aes(x=pl_orbper,y=pl_Rforecast2),
                                  shape=21,fill=mycol3,col="black", size=2)
postplot <- postplot + geom_point(data=postkepler,aes(x=pl_orbper,y=pl_rade),
                                 shape=21,fill=mycol2,col="black",size=2)


print(postplot)
ggsave("plots/post_kepler.eps")

keplerlist <- c("Kepler","K2")
kepler<-filter(frankenplanets,pl_facility %in% keplerlist)

postplot2 <- preplot + geom_point(data=kepler,aes(x=pl_orbper,y=pl_rade),
                                 shape=21,fill=mycol2,col="black",size=2)

print(postplot2)
ggsave("plots/kepleronly.eps")


notkep_post <- filter(postkepler,pl_facility != "Kepler")

notkeplerplot<- preplot + geom_point(data=notkep_post,aes(x=pl_orbper,y=pl_rade),
                                 shape=21,fill=mycol5,col="black",size=2) 
notkeplerplot <- notkeplerplot + geom_point(data=postkepler_rv,aes(x=pl_orbper,y=pl_Rforecast2),
                                  shape=21,fill=mycol5,col="black", size=2)
print(notkeplerplot)
ggsave("plots/notkepler.eps")

pastaplot<- notkeplerplot + geom_point(data=kepler,aes(x=pl_orbper,y=pl_rade),
                                 shape=21,fill=mycol2,col="black",size=2)
pastaplot <- pastaplot + geom_point(data=k2,aes(x=pl_orbper,y=pl_rade),
                                 shape=21,fill=mycol2,col="black",size=2)

print(pastaplot)
ggsave("plots/post_kepler2.eps")
```


Maybe add one more historical plot.  I'd like to show which fraction of planets came from Kepler.  I'm thinking a cumulative plot is the best method for this...

```{r}
# first let's add a new column that is either detection method or Kepler

keplerlist <- c("Kepler", "K2")

frankenplanets <- mutate(frankenplanets, mymethod = ifelse(pl_facility %in% keplerlist, "Kepler", pl_discmethod))


```

```{r}
baseplot <- ggplot(frankenplanets)
pp1 <- baseplot + stat_ecdf(aes(pl_disc),geom = "area", pad = FALSE)
print(pp1)

pp1 <- baseplot + stat_ecdf(aes(pl_disc,fill=mymethod),geom="area")
print(pp1)
```

Not quite what I wanted.  Let's try this...

```{r}

frankenplanets$dummy <- rep(1,nrow(frankenplanets))

baseplot <- ggplot(data = frankenplanets %>% group_by(pl_disc) %>% summarise(count=n()),aes(x=pl_disc,y=count))
pp1 <- baseplot + geom_area(aes(y=cumsum(count)))
print(pp1)

kep <- frankenplanets %>% filter(mymethod == "Kepler")
pp2 <- pp1 + geom_area(data=kep %>% group_by(pl_disc) %>% summarise(count=n()),
                       aes(x=pl_disc,y=cumsum(count)),fill="blue")
print(pp2)


```

ggplot(NULL,aes(x))+geom_histogram(aes(y=cumsum(..count..)))+
       stat_bin(aes(y=cumsum(..count..)),geom="line",color="green")


```{r}

baseplot <- ggplot(frankenplanets) + theme_bw()
pp1 <- baseplot + geom_histogram(aes(x=pl_disc,y=cumsum(..count..)))
print(pp1)

kep <- frankenplanets %>% filter(mymethod == "Kepler")
pp2 <- pp1 + geom_histogram(data=kep,aes(x=pl_disc,y=cumsum(..count..)),fill="blue")
print(pp2)

```


```{r}

mycol1<-"blue"
mycol2<-"yellow"
mycol5<-"pink"

baseplot <- ggplot(frankenplanets) + theme_kepler() + ylab("Cummulative # of Planets") + xlab("Year")
baseplot <- baseplot + xlim(c(2000, 2020))
pp1 <- baseplot + geom_bar(aes(x=pl_disc,y=cumsum(..count..)),fill=mycol1)
print(pp1)

kep <- frankenplanets %>% filter(mymethod == "Kepler")
pp2 <- pp1 + geom_bar(data=kep,aes(x=pl_disc,y=cumsum(..count..)),fill=mycol2)
print(pp2)

ggsave("plots/cum_counts.eps")


baseplot <- ggplot(frankenplanets) + theme_kepler() + ylab("Cummulative # of Planets") + xlab("Year")
baseplot <- baseplot + xlim(c(2000, 2020))
pp1 <- baseplot + geom_bar(aes(x=pl_disc,y=cumsum(..count..)),fill=mycol5)
print(pp1)

kep <- frankenplanets %>% filter(mymethod == "Kepler")
pp2 <- pp1 + geom_bar(data=kep,aes(x=pl_disc,y=cumsum(..count..)),fill=mycol2)
print(pp2)

ggsave("plots/cum_counts2.eps")

```



